State and Parameter Estimation for a Spatio-Temporal Neural Fields

### 2010-04-14 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
changing some parameter names so we can use UKF literature if we need to cut it back

\section{Neural Field Model}
In this section, we describe a variant of the  Wilson and Cowan neural field model~\cite{Wilson1973} that will be used to create our intracortical connectivity estimator. This model is descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behaviour. 
\begin{todo}
	{Really need to add some citations here.}
\end{todo}
It is also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations. 
\begin{todo}
	{Same here.}
\end{todo}
The neural field model is popular due to being parsimonious yet having a strong link with the underlying physiology. The model relates the average number of action potentials $g(\mathbf{r},t)$ arriving at position $r$ to the local post synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials generated at a neuronal population at location $\mathbf{r}$ by action potentials arriving from all other connected populations at locations $\mathbf{r}'$ can be described by
\begin{equation}\label{SpikesToPotential}
	v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'}.
\end{equation}
The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{SynapticRespKernel}
	h(t) = \eta(t)\exp{\left(-\alpha t\right)}.
\end{equation}
where $\alpha=\tau^{-1}$ and $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{RateBasedInteractions}
	g\left( \mathbf{r},t \right) = \int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)d\mathbf{r}'} + u\left(\mathbf{r},t\right),
\end{equation}
where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface. The term $u\left(\mathbf{r},t\right)$ represents a known external input. The connectivity kernel is typically a ``Mexican hat'' function \todo{which describes strong local activation, weak mid-range repression and weak long-range activation \cite{}}. The exact shape of this kernel is assumed to vary across patients, and hence needs to be inferred from data.

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
	f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \beta \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}.
\end{equation}
%Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. 
By substituting \ref{RateBasedInteractions} into \ref{SpikesToPotential}  we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
	v\left(\mathbf{r},t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega   w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'.
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
	Lh\left( t \right) = \delta \left( t \right),
\end{equation}
\todo{Is the above right? I can't think of any way to make $\frac{d}{dt} h(t) + \alpha h(t)$ equal to $\delta(t)$}
where $L=\frac{d}{dt} + \alpha$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}\label{FinalFormContinuous}
	\frac{dv\left( \mathbf{r},t \right)}{dt} + \alpha v\left( \mathbf{r},t \right) = \int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)dr'} + u\left(\mathbf{r},t\right).
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model we discretize time using a first-order Euler method (see Appendix~\ref{Time Discretization}) giving
\begin{equation}\label{DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = \lambda v_t\left(\mathbf{r}\right) + T \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right),
\end{equation} 
where $T$ is the time step, $\lambda = 1-\alpha T$ and is an iid spatially colored, temporally white disturbance term with the covariance 
\begin{equation}
    cov(\mathbf{e}_t,\mathbf{e}_{t+\tau}) = \left\{ \begin{array}{*{20}{c}}
    \gamma\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
    0 & \mathrm{otherwise}. \\
    \end{array} \right.
\end{equation}
This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t,
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\varepsilon_t \sim \mathcal{N}\left(0,\Sigma{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}.
\end{equation}

@@

	\boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T\boldsymbol{\theta}^{\top}\int_\Omega  {\boldsymbol{\psi} \left(\mathbf{r}-\mathbf{r}'\right) f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)d\mathbf{r}'}  - \lambda\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right).

@@

	\mathbf{x}_t &=& q(\mathbf{x}_t)\boldsymbol{\theta} - \lambda\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\

@@

\boldsymbol{\Sigma}_e=\boldsymbol{\Gamma}^{-1}\int\limits_{\Omega}\int\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \rho\left(\mathbf{r}- \mathbf{r}' \right)\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf{r}\boldsymbol{\Gamma}^{- \top} 

@@

Due to the high dimensionality of the brain and the current electrode systems, we can not expect to have more sensors than basis functions.
\section{State and Parameter Estimation}
The estimation process is a two part iterative algorithm that yields the estimate of the connectivity kernel parameters and the posterior distribution of the field weights. In the parameter estimation step, Least Squares (LS) method is used which minimizes the sum of the squares of the errors on each state equation (see Appendix). In the state estimation step, given an estimate of $\boldsymbol \theta$ and $\lambda$ the Unscented Rauch Tung Striebel Smoother (URTSS) \cite{Sarkka2010} was used to determine the posterior state sequence.
The (URTSS) uses the Unscented Kalman Filter (UKF) in the forward iteration \cite{Julier1997} along with a separate backward smoothing path to compute the smooth estimations. The method is appropriate for state estimation of stochastic nonlinear dynamical systems. Alternatives to the UKF are the Extended Kalman Filter (EKF) \cite{Haykin2001}  and Sequential Monte Carlo (SMC) filtering \cite{doucet2001}. The extended Kalman filter approximates the state transition equation by linearizing about the fixed points (by calculating the Jacobian) at each time step. The linearization maintains the Gaussianality in the model. The unscented Kalman filter approximates the \emph{a posteriori} state density by a Gaussian distribution using a minimal set of carefully chosen sigma points, whilst maintaining the nonlinearity in the system. This has been shown to give superior performance over the EKF in state estimation, as the EKF maintains a first order approximation where the UKF provides an approximation accurate at least to the second order. In addition, the computation of the Jacobian used in the
EKF can be problematic. SMC filtering can theoretically provide an exact \emph{a posteriori} state density for nonlinear systems. However, currently this method is not appropriate for our problem, due to computation demands.

@@

 The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t)\boldsymbol{\theta} - \lambda\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\lambda$ are the systems parameters. In forwards iteration since the observation equation is linear in parameters we only need to use Unsented transformation in prediction step. The update step equations remains the same as the standard Kalman filtering formulations. In backward smoothing path, estimation from the forward filtering is used to form the new set of sigma points in a similar way to the filterin step, however the predicted cross-covariance of the states need to be calculated. A summary of the algorithm is given in Algorithm 
\begin{algorithm}
\caption{The Unscented Kalman Filter}\label{CPTEMDAlgorithm}
\begin{algorithmic}[1]
\State Initialise
\begin{equation*}
\hat{\mathbf x}_0, \mathbf P_0
\end{equation*}
\State Forward iteration
For $t \in \left\lbrace 0,\cdots, N\right\rbrace $,
\begin{equation*}
\mathcal X_t^f=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^{f}-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
Propogate the sigma points through the dynamic model
\begin{equation}
\mathcal X_{i,t+1|t}^{f*}=Q(\mathcal X_t^f,u_t)
\end{equation}
\begin{equation*}
 \hat{\mathbf x}_{k +t}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1|t}^{f*}
\end{equation*}
\begin{equation*}
 \mathbf P_{t +1}^-=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e
\end{equation*}
\begin{equation*}
\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_v)^{-1}
\end{equation*}
\begin{equation*}
\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-})
\end{equation*}
\begin{equation*}
 \mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}
\end{equation*}
\State Backward iteration
\State initialisation
\begin{equation}
 \mathbf P_T^s= \mathbf P_T^f
\end{equation}
\begin{equation}
\hat{\mathbf x}^s_T= \hat{\mathbf x}^f_T
\end{equation}
For $t \in \left\lbrace N-1, \cdots, 0 \right\rbrace $
\begin{equation*}
\mathcal X_t^s=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^f-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
\end{algorithmic}
\end{algorithm}
\section{Results}

@@

\frac{v\left( \mathbf{r},t+T \right) - v\left( \mathbf{r},t\right)}{T} =   -\alpha v\left( \mathbf{r},t \right) + \int_\Omega  {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'} + u\left(\mathbf{r},t\right).

@@

v_{t+1}\left( \mathbf{r}\right) &=& v_t\left( \mathbf{r}\right) -T \alpha v_t\left( \mathbf{r}\right) + T \int_\Omega  {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'} + T u_t\left(\mathbf{r}\right).

@@

	v_{t+1}\left(\mathbf{r}\right) = \lambda v_t\left(\mathbf{r}\right) + T \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T u_t\left(\mathbf{r}\right),

@@

where $\lambda = 1 - T \alpha$.

@@

	v_{t+1}\left(\mathbf{r}_{ij}\right) = \lambda v_t\left(\mathbf{r}_{ij}\right) + T \Delta^2 \sum_{i=1}^{i=I}{ \sum_{j=1}^{j=J}{ w\left( \mathbf{r}-\mathbf{r}_{ij}' \right)f\left( v_t\left( \mathbf{r}_{ij}'\right) \right)} } + T u_t\left(\mathbf{r}\right) + e_t(\mathbf{r}_{i,j}),

@@

	\int_\Omega  {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) d\mathbf{r}} \mathbf{x}_{t+1} = T \int_\Omega  {\boldsymbol{\phi} (\mathbf{r}) \boldsymbol{\theta}^{\top} \int_\Omega  {\boldsymbol{\psi}  \left(\mathbf{r}-\mathbf{r}'\right) f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right) \mathbf{x}_t \right)d\mathbf{r}'}d\mathbf{r}} - \lambda\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\phi}^{\top}(\mathbf{r})d\mathbf{r}} \mathbf{x}_t \\

@@

	\mathbf{x}_{t+1} = T\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right) \boldsymbol{\psi}^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \lambda \mathbf{x}_t \\

@@

\boldsymbol\Sigma_e =\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \beta\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top} 

@@

\subsection{proof}

@@

 \mathbf x_{t+1}=\mathbf{x}_{t}+T_s \mathbf{Q}(\mathbf{x}_t)\boldsymbol{\theta}-\alpha T_s\mathbf{x}_t+T_s\Gamma^{-1}\int_\Omega\boldsymbol{\phi}(\mathbf r)\varepsilon_t(\mathbf r)d\mathbf r

@@

 \mathbf x_{1}&=&\mathbf x_{0}+T_s \mathbf Q(\mathbf x_0) \boldsymbol{\theta}-\alpha T_s\mathbf x_0+T_s\mathbf e_0 \nonumber \\
 \mathbf x_{2}&=&\mathbf x_{1}+T_s \mathbf Q(\mathbf x_1) \boldsymbol{\theta}-\alpha T_s\mathbf x_1+T_s\mathbf e_1\nonumber\\

@@

 \mathbf x_{n}&=&\mathbf x_{n-1}+T_s \mathbf Q(\mathbf x_{n-1}) \boldsymbol{\theta}-\alpha T_s\mathbf x_{n-1}+T_s\mathbf e_{n-1}

@@

	\left[\begin{array}{cc} \boldsymbol{\theta} \\ \alpha\end{array}\right]+T_s

@@

,\quad \mathbf W=\left[\begin{array}{cc} \boldsymbol{\theta} \\ \alpha\end{array}\right],\quad \boldsymbol \xi=T_s\left[\begin{array}{cccc}\mathbf e_0\\\mathbf e_1\\\vdots\\\mathbf e_{n-1}\end{array}\right]

>>>>

\section{Neural Field Model}
In this section, we describe a variant of the  Wilson and Cowan neural field model~\cite{Wilson1973} that will be used to create our intracortical connectivity estimator. This model is descriptive of a range of neurodynamics of the cortex such as evoked potentials, visual hallucinations, and epileptic behaviour. 
\begin{todo}
	{Really need to add some citations here.}
\end{todo}
It is also capable of generating complex patterns of activity such as Turing patterns, spirals, and traveling oscillations. 
\begin{todo}
	{Same here.}
\end{todo}
The neural field model is popular due to being parsimonious yet having a strong link with the underlying physiology. The model relates the average number of action potentials $g(\mathbf{r},t)$ arriving at position $r$ to the local post synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials generated at a neuronal population at location $\mathbf{r}$ by action potentials arriving from all other connected populations at locations $\mathbf{r}'$ can be described by
\begin{equation}\label{SpikesToPotential}
	v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'}.
\end{equation}
The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{SynapticRespKernel}
	h(t) = \eta(t)\exp{\left(-\zeta t\right)}.
\end{equation}
where $\zeta=\tau^{-1}$ and $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{RateBasedInteractions}
	g\left( \mathbf{r},t \right) = \int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)d\mathbf{r}'} + u\left(\mathbf{r},t\right),
\end{equation}
where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface. The term $u\left(\mathbf{r},t\right)$ represents a known external input. The connectivity kernel is typically a ``Mexican hat'' function \todo{which describes strong local activation, weak mid-range repression and weak long-range activation \cite{}}. The exact shape of this kernel is assumed to vary across patients, and hence needs to be inferred from data.

The firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
	f\left( v\left( \mathbf{r}', t \right) \right) = \frac{f_{max}\left(\mathbf{r}\right)}{1 + \exp \left( \varsigma \left( v_0\left( \mathbf{r} \right) - v\left(\mathbf{r}',t\right) \right) \right)}.
\end{equation}
%Even though we will model $v_0$ and $\nu_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality at this point. 
By substituting \ref{RateBasedInteractions} into \ref{SpikesToPotential}  we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
	v\left(\mathbf{r},t\right) = \int_{-\infty}^t h\left(t - t'\right) \left(\int_\Omega   w\left(\mathbf{r},\mathbf{r}'\right) f\left( v\left( \mathbf{r}',t \right)\right)d\mathbf{r}' + u(\mathbf{r},t) \right)dt'.
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
	Lh\left( t \right) = \delta \left( t \right),
\end{equation}
where $L=\frac{d}{dt} + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function giving 
\begin{equation}\label{FinalFormContinuous}
	\frac{dv\left( \mathbf{r},t \right)}{dt} + \zeta v\left( \mathbf{r},t \right) = \int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)dr'} + u\left(\mathbf{r},t\right).
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model we discretize time using a first-order Euler method (see Appendix~\ref{Time Discretization}) giving
\begin{equation}\label{DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right),
\end{equation} 
where $T$ is the time step, $\xi = 1-\zeta T$ and is an iid spatially colored, temporally white disturbance term with the covariance 
\begin{equation}
    cov(\mathbf{e}_t,\mathbf{e}_{t+\tau}) = \left\{ \begin{array}{*{20}{c}}
    \gamma\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
    0 & \mathrm{otherwise}. \\
    \end{array} \right.
\end{equation}
This term was added to drive the system and account for uncertainty in the model. To generate iEEG/LFP data we used the output function 
\begin{equation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)d\mathbf{r}'} + \boldsymbol{\varepsilon}_t,
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field where $n=1,...,N$ indexes the sensors and $\varepsilon_t \sim \mathcal{N}\left(0,\Sigma{\varepsilon}\right)$. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the geometry of pick-up range of sensors where
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)}.
\end{equation}

@@

	\boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T\boldsymbol{\theta}^{\top}\int_\Omega  {\boldsymbol{\psi} \left(\mathbf{r}-\mathbf{r}'\right) f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)d\mathbf{r}'}  - \xi\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right).

@@

	\mathbf{x}_t &=& q(\mathbf{x}_t)\boldsymbol{\theta} - \xi\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\

@@

\boldsymbol{\Sigma}_e=\boldsymbol{\Gamma}^{-1}\int\limits_{\Omega}\int\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf{r}- \mathbf{r}' \right)\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf{r}\boldsymbol{\Gamma}^{- \top} 

@@

Due to the high dimensionality of the brain and the current electrode systems, we can not expect to have more sensors than basis functions.
\section{State and Parameter Estimation}
The estimation process is a two part iterative algorithm that yields the estimate of the connectivity kernel parameters and the posterior distribution of the field weights. In the parameter estimation step, Least Squares (LS) method is used which minimizes the sum of the squares of the errors on each state equation (see Appendix). In the state estimation step, given an estimate of $\boldsymbol \theta$ and $\xi$ the Unscented Rauch Tung Striebel Smoother (URTSS) \cite{Sarkka2010} was used to determine the posterior state sequence.
The (URTSS) uses the Unscented Kalman Filter (UKF) in the forward iteration \cite{Julier1997} along with a separate backward smoothing path to compute the smooth estimations. The method is appropriate for state estimation of stochastic nonlinear dynamical systems. Alternatives to the UKF are the Extended Kalman Filter (EKF) \cite{Haykin2001}  and Sequential Monte Carlo (SMC) filtering \cite{doucet2001}. The extended Kalman filter approximates the state transition equation by linearizing about the fixed points (by calculating the Jacobian) at each time step. The linearization maintains the Gaussianality in the model. The unscented Kalman filter approximates the \emph{a posteriori} state density by a Gaussian distribution using a minimal set of carefully chosen sigma points, whilst maintaining the nonlinearity in the system. This has been shown to give superior performance over the EKF in state estimation, as the EKF maintains a first order approximation where the UKF provides an approximation accurate at least to the second order. In addition, the computation of the Jacobian used in the
EKF can be problematic. SMC filtering can theoretically provide an exact \emph{a posteriori} state density for nonlinear systems. However, currently this method is not appropriate for our problem, due to computation demands.

@@

 The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t)\boldsymbol{\theta} - \xi\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\xi$ are the systems parameters. In forwards iteration since the observation equation is linear in parameters we only need to use Unsented transformation in prediction step. The update step equations remains the same as the standard Kalman filtering formulations. In backward smoothing path, estimation from the forward filtering is used to form the new set of sigma points in a similar way to the filterin step, however the predicted cross-covariance of the states need to be calculated. A summary of the algorithm is given in Algorithm 
\begin{algorithm}
\caption{The Unscented Kalman Filter}\label{CPTEMDAlgorithm}
\begin{algorithmic}[1]
\State Initialise
\begin{equation*}
\hat{\mathbf x}_0, \mathbf P_0
\end{equation*}
\State Forward iteration
For $t \in \left\lbrace 0,\cdots, N\right\rbrace $,
\begin{equation*}
\mathcal X_t^f=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^{f}-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
Propogate the sigma points through the dynamic model
\begin{equation}
\mathcal X_{i,t+1|t}^{f*}=Q(\mathcal X_t^f,u_t)
\end{equation}
\begin{equation*}
 \hat{\mathbf x}_{k +t}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1|t}^{f*}
\end{equation*}
\begin{equation*}
 \mathbf P_{t +1}^-=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e
\end{equation*}
\begin{equation*}
\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_v)^{-1}
\end{equation*}
\begin{equation*}
\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-})
\end{equation*}
\begin{equation*}
 \mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}
\end{equation*}
\State Backward iteration
\State initialisation
\begin{equation}
 \mathbf P_T^s= \mathbf P_T^f
\end{equation}
\begin{equation}
\hat{\mathbf x}^s_T= \hat{\mathbf x}^f_T
\end{equation}
For $t \in \left\lbrace N-1, \cdots, 0 \right\rbrace $
\begin{equation*}
\mathcal X_t^s=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^f-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
\end{algorithmic}
\end{algorithm}
\section{Results}

@@

\frac{v\left( \mathbf{r},t+T \right) - v\left( \mathbf{r},t\right)}{T} =   -\zeta v\left( \mathbf{r},t \right) + \int_\Omega  {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'} + u\left(\mathbf{r},t\right).

@@

v_{t+1}\left( \mathbf{r}\right) &=& v_t\left( \mathbf{r}\right) -T \zeta v_t\left( \mathbf{r}\right) + T \int_\Omega  {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'} + T u_t\left(\mathbf{r}\right).

@@

	v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T u_t\left(\mathbf{r}\right),

@@

where $\xi = 1 - T \zeta$.

@@

	v_{t+1}\left(\mathbf{r}_{ij}\right) = \xi v_t\left(\mathbf{r}_{ij}\right) + T \Delta^2 \sum_{i=1}^{i=I}{ \sum_{j=1}^{j=J}{ w\left( \mathbf{r}-\mathbf{r}_{ij}' \right)f\left( v_t\left( \mathbf{r}_{ij}'\right) \right)} } + T u_t\left(\mathbf{r}\right) + e_t(\mathbf{r}_{i,j}),

@@

	\int_\Omega  {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) d\mathbf{r}} \mathbf{x}_{t+1} = T \int_\Omega  {\boldsymbol{\phi} (\mathbf{r}) \boldsymbol{\theta}^{\top} \int_\Omega  {\boldsymbol{\psi}  \left(\mathbf{r}-\mathbf{r}'\right) f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right) \mathbf{x}_t \right)d\mathbf{r}'}d\mathbf{r}} - \xi\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\phi}^{\top}(\mathbf{r})d\mathbf{r}} \mathbf{x}_t \\

@@

	\mathbf{x}_{t+1} = T\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right) \boldsymbol{\psi}^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \xi \mathbf{x}_t \\

@@

\boldsymbol\Sigma_e =\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top} 

@@

\subsection*{Proof}

@@

 \mathbf x_{t+1}=\mathbf{x}_{t}+T_s \mathbf{Q}(\mathbf{x}_t)\boldsymbol{\theta}-\zeta T_s\mathbf{x}_t+T_s\Gamma^{-1}\int_\Omega\boldsymbol{\phi}(\mathbf r)\varepsilon_t(\mathbf r)d\mathbf r

@@

 \mathbf x_{1}&=&\mathbf x_{0}+T_s \mathbf Q(\mathbf x_0) \boldsymbol{\theta}-\zeta T_s\mathbf x_0+T_s\mathbf e_0 \nonumber \\
 \mathbf x_{2}&=&\mathbf x_{1}+T_s \mathbf Q(\mathbf x_1) \boldsymbol{\theta}-\zeta T_s\mathbf x_1+T_s\mathbf e_1\nonumber\\

@@

 \mathbf x_{n}&=&\mathbf x_{n-1}+T_s \mathbf Q(\mathbf x_{n-1}) \boldsymbol{\theta}-\zeta T_s\mathbf x_{n-1}+T_s\mathbf e_{n-1}

@@

	\left[\begin{array}{cc} \boldsymbol{\theta} \\ \zeta\end{array}\right]+T_s

@@

,\quad \mathbf W=\left[\begin{array}{cc} \boldsymbol{\theta} \\ \zeta\end{array}\right],\quad \boldsymbol \xi=T_s\left[\begin{array}{cccc}\mathbf e_0\\\mathbf e_1\\\vdots\\\mathbf e_{n-1}\end{array}\right]
